---
title: "Report for Apprehensions at the US-Mexico border from 2000 to 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1.Introduction

## 2.Comparation between 2010 and 2017
```{r}
#function that turn the sector col to row name
rowname <- function(x){
  rownames(x) <- x[,1]
  x<-subset(x,select=-Sector)
  x
}

#load the data and clean them
year2010 <- read.csv("BP Apprehensions 2010.csv")
year2017 <- read.csv("PB Apprehensions 2017.csv")
year2010 <- rowname(year2010)
year2017 <- rowname(year2017)
sector <- rownames(year2010)
month <- colnames(year2010)

#function that shows side by side graph for all sector

#function that shows side by side graph for each month

```

##2.1
## 3.The statistical tests between 2010 and 2017

### T-test Results and Interpretations

#### 3.1 T-test between the sector with the most apprehensions in 2010 and 2017
```{r pressure, echo=T}
#load the data
year2010 <- read.csv("bp appre 2010.csv")
year2017 <- read.csv("bp appre 2017.csv")

#function that turn the sector col to row name
operation_row <- function(x){
  rownames(x) <- x[,1]
  x<-subset(x,select=-Sector)
  x <- rbind(x, colSums(x))
  -length(rownames(x))  
  rownames(x) <- c(rownames(x)[-length(rownames(x))], "Total")
  x <- cbind(x,rowSums(x))
  colnames(x) <- c(colnames(x)[-length(colnames(x))], "Total")
  x
}

#clean the data
year2010 <- operation_row(year2010)
year2017 <- operation_row(year2017)

#select the sector with the most apprehensions in 2010 and 2017
year2010_rowtotal<-year2010[-nrow(year2010),]
year2017_rowtotal<-year2017[-nrow(year2017),]
most_sec_2010<-(year2010_rowtotal[year2010_rowtotal$Total==max(year2010_rowtotal$Total),])
most_sec_2017<-(year2017_rowtotal[year2017_rowtotal$Total==max(year2017_rowtotal$Total),])
most_sec_2010<-most_sec_2010[,-ncol(most_sec_2010)]
most_sec_2017<-most_sec_2017[,-ncol(most_sec_2017)]

#do the t-test, compare the sector with the most apprehensions in two years
most_sec_2010<-t(most_sec_2010)
most_sec_2017<-t(most_sec_2017)
print(paste(colnames(most_sec_2010),"is sector with most apprehension in 2010")) 
print(paste(colnames(most_sec_2017),"is sector with most apprehension in 2017")) 
t.test(most_sec_2010,most_sec_2017,paired=T)
```
t test result: p= 0.1061 > 0.05  

Explaination: We know that the sector with most apprehension in 2010 is Tucson and the sector with most apprehension in 2017 is Rio Grande Valley. From the resulr of t-test, We cannot reject the null hypothesis that the apprehension in Tucson in 2010 has equal estimated means as Rio Grande Valley in 2017. Though the sector with most apprehension changed, but the number of most apprehension doesn't change.

#### 3.2 T-test between the 3 month periods with the most apprehensions in 2010 and 2017
```{r}
#select the 3 month periods with the most apprehensions in 2010 and 2017
year2010_coltotal <- year2010[,-ncol(year2010)]
year2017_coltotal <- year2017[,-ncol(year2010)]

#sort the data in 2010,2017 with respect to total apprehension by month
year2010_coltotal <- year2010_coltotal[,order(year2010_coltotal["Total",],decreasing = T)]
year2017_coltotal <- year2017_coltotal[,order(year2017_coltotal["Total",],decreasing = T)]

#select the months with top three total apprehension
most_month_2010 <-  year2010_coltotal[,1:3]
most_month_2017 <-  year2017_coltotal[,1:3]
most_month_2010 <- most_month_2010[-nrow(most_month_2010),]
most_month_2017 <- most_month_2017[-nrow(most_month_2017),]


a <- cbind(most_month_2010[,1],most_month_2010[,2],most_month_2010[,3])
b <- cbind(most_month_2017[,1],most_month_2017[,2],most_month_2017[,3])
#do the t-test, compare the sector with the most apprehensions in two years
print(paste(colnames(most_month_2010),
            "is one of the 3 month periods with the most apprehensions in 2010")) 
print(paste(colnames(most_month_2017),
            "is one of the 3 month periods with the most apprehensions in 2017")) 
t.test(a,b,paired=T)
```
t test result: p= 0.5904 > 0.05  

Explaination: We know that the 3 month periods with the most apprehensions in 2010 is March, April and May, and the 3 month periods with the most apprehensions in 2017 is November, October, December. From the resulr of t-test, We cannot reject the null hypothesis that the apprehension in March, April and May 2010 has equal estimated means as November, October, December in 2017. Though the 3 months with most apprehension changed, but the number of most apprehension doesn't change to o much.

## 4.The apprehensions change from 2000 to 2017
```{r}
#load, clean and rerrange the data
ts<-read.csv("PB monthly summaries.csv")
rownames(ts) <- ts[,1]
ts <- ts[order(ts$year),]
ts<-subset(ts,select=-year)

#turn it into timeseries format
ts_df<-ts
ts <- as.vector(t(ts))
ts<-ts(ts,start = c(2000,1), frequency=12)

#plot the timeseries data
ts.plot(ts, 
        col = 3, 
        xlab="year", 
        ylab="Apprehensions", 
        lty=c(1:3), 
        main ="Apprehension from 2000 to 2017")

#draw the avg
ts_avg<-apply(ts_df,1,mean)
ts_avg<-ts(ts_avg,start = c(2000), frequency=1)
points(ts_avg,col = 4,pch=20)
lines(ts_avg,col = 4,pch=20,lty=2)

#get the max/min month index
pos_max<-which(ts_df == max(ts_df), arr.ind=T)
pos_min<-which(ts_df == min(ts_df), arr.ind=T)

#get the month vector
month<-c("October","November","December","January","February","March","April","May","June","July","August","September")

#get the exact time point of the maximum/minimum element in timeseries data
max_index <- time(ts)[ts==max(ts)]
min_index <- time(ts)[ts==min(ts)]
max_avg <- time(ts_avg)[ts_avg==max(ts_avg)]
min_avg <- time(ts_avg)[ts_avg==min(ts_avg)]

#label the maximum/minimum apprehension in the graph
text(max_index,max(ts),
     paste("num:",max(ts),"\n","date:",floor(max_index),month[as.vector(pos_max)[2]]),
     cex=0.4,pos=4)
text(min_index,min(ts),
     paste("num:",min(ts),"\n","date:",floor(min_index),month[as.vector(pos_min)[2]]),
     cex=0.4,pos=2)
text(max_avg,
     max(ts_avg),
     paste("avg max:",floor(max(ts_avg)),"\n","year:",floor(max_avg)),
     cex=0.4,
     pos=4)
text(min_avg,
     min(ts_avg),
     paste("avg min:",floor(min(ts_avg)),"\n","year:",floor(min_avg)),
     cex=0.4,
     pos=3)
text(2005,
     ts_avg[time(ts_avg)==2005],
     paste("avg:",floor(ts_avg[time(ts_avg)==2005]),"\n","year:",2005),
     cex=0.4,
     pos=3)

#pinpoint the maximum/minimum apprehension in the graph
points(max_index,max(ts),pch=19,col=2)
points(min_index,min(ts),pch=19,col=7)

#add a legend which denotes the meaning of the points
legend("topright", 
       c("maximum apprehension",
         "minimum apprehension",
         "annual average apprehension"),
       col = c(2,7,4),
       pch =c(19,19,20),
       lty=c(0,0,2),
       bty = "n",
       bg = 'gray90')
```
  As what is shown in the graph above. Apprehensions at the US-Mexico border reached its historic lows in the 2017 and shows a downward trend since 2000. Dropping a the way from its peak of 220063 in 2000 to its bottom as 11127, nearly 2000% decline. 
  
  As for the annual average apprehension, we can see it has been dropped from 136973 in 2000 to 25326 in 2017. Though raised a little bit in 2005 to 97616, it went all the way down to 25326 in 2017.
  
  



