---
title: "Report for Apprehensions at the US-Mexico border from 2000 to 2017"
author: "Hongyi He, Glib Dolotov"
date: "February 19, 2018"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1.Introduction

## 2.Comparation between 2010 and 2017

```{r, code=readLines("part1_comparison.R")}
```

##2.1
## 3.The statistical tests between 2010 and 2017

### T-test Results and Interpretations

#### 3.1 T-test between the sector with the most apprehensions in 2010 and 2017
```{r pressure, code = readLines("part2_t-test.R")[1:34]}
```
t test result: p= 0.1061 > 0.05  

Explaination: We know that the sector with most apprehension in 2010 is Tucson and the sector with most apprehension in 2017 is Rio Grande Valley. From the resulr of t-test, We cannot reject the null hypothesis that the apprehension in Tucson in 2010 has equal estimated means as Rio Grande Valley in 2017. Though the sector with most apprehension changed, but the number of most apprehension doesn't change.

#### 3.2 T-test between the 3 month periods with the most apprehensions in 2010 and 2017
```{r, code = readLines("part2_t-test.R")[36:58]}
```
t test result: p= 0.5904 > 0.05  

Explaination: We know that the 3 month periods with the most apprehensions in 2010 is March, April and May, and the 3 month periods with the most apprehensions in 2017 is November, October, December. From the resulr of t-test, We cannot reject the null hypothesis that the apprehension in March, April and May 2010 has equal estimated means as November, October, December in 2017. Though the 3 months with most apprehension changed, but the number of most apprehension doesn't change to o much.

## 4.The apprehensions change from 2000 to 2017
```{r, code = readLines("part3_timeseries.R")}
```
  As what is shown in the graph above. Apprehensions at the US-Mexico border reached its historic lows in the 2017 and shows a downward trend since 2000. Dropping a the way from its peak of 220063 in 2000 to its bottom as 11127, nearly 2000% decline. 
  
  As for the annual average apprehension, we can see it has been dropped from 136973 in 2000 to 25326 in 2017. Though raised a little bit in 2005 to 97616, it went all the way down to 25326 in 2017.
  
  



